---

---

## Data

Consistency is key when managing your data!

1. At a minimum, use the same terms to refer to the same things throughout your data.

2. Potentially use a discipline-wide published vocabulary (ex: [hydrological controlled vocab.](http://vocabulary.odm2.org/))

===

### Naming data files - review from basic-R-lesson

Bad:
"CeNsus data*2ndTry 2/15/2017.csv"   

Good:
"census_data_try_2.csv"  

```{r}
library(tidyverse)

stm_dat <- read_csv("data/data_package/StormEvents.csv")
```

===

### Formatting data - review from data-manipulation-R-lesson

1. One table/file for each type of observation.
2. Each variable has its own column.
3. Each observation has its own row.
4. Each value has its own cell.

Downstream operations/analysis require tidy data. You can work towards read-to-analyze data incrementally,
documenting the intermediate data and steps you took in your scripts.  This can be a powerful accelerator 
for your analysis both now and in the future.  

===

Let's do a few checks to see if our data is tidy.

 - make sure data meet criteria above
 - make sure blanks are NA or some other standard
```{r}
head(stm_dat)
tail(stm_dat)
```    
 
 - check date format
```{r}
str(stm_dat)    
```  

 - check case, white space, etc.  
```{r}
unique(stm_dat$EVENT_NARRATIVE)    
```
    
===
    
### Outputting derived data - review from the basic-R-lesson
    
Because we followed the principle of one table/file for each type of observation, 
we have one table to be written out to a file now.  

```{r}
write_csv(stm_dat, "data_package/StormEvents_d2006.csv")
```

===
    
### Versioning your derived data

Issue: Making sure you and your collaborators are using the same version of a dataset

One solution: version your data in a repository
    Upload a version of your data early on in your research data process.
    Versioning in the repository will assign a unique ID to each version of your data
    Can download/use a particular version of the data (even if not the most recent version)
    Tell collaborators which version you used for analysis
    
May help address the issue of collaborators using different versions of the data for analysis.


